---
title: "How to Create an Interactive WebGL Network Graph Using R and Plotly"
author: "Max Woolf (@minimaxir)"
date: "December 5th, 2016"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
---

This R Notebook is the complement to my blog post [How to Create an Interactive WebGL Network Graph Using R](http://minimaxir.com/2016/12/interactive-network/).

This notebook is licensed under the MIT License. If you use the code or data visualization designs contained within this notebook, it would be greatly appreciated if proper attribution is given back to this notebook and/or myself. Thanks! :)

# Setup

Setup the R packages.

```{r}

# must install ggnetwork using from source to avoid ggplot2 2.2.0 issue
# install.packages("ggnetwork", type="source")

library(dplyr)
library(readr)
library(igraph)
library(intergraph)
library(sna)
library(ggplot2)
library(ggnetwork)
library(plotly)
library(htmlwidgets)

sessionInfo()
```

```{r}
df <- read_csv('hibp_edges.csv')
df %>% arrange(desc(Weight)) %>% head()
```

There are `r df %>% nrow()` edges.

```{r}
net <- graph.data.frame(df, directed = T)

V(net)$degree <- centralization.degree(net)$res
V(net)$weighted_degree <- graph.strength(net, weights=V(net)$Weight)

net
```

```{r}
V(net)$group <- membership(cluster_walktrap(net, weights=E(net)$Weight))
V(net)$centrality <- eigen_centrality(net, weights=E(net)$Weight)$vector
```

```{r}
net <- igraph::delete.vertices(net, V(net)[weighted_degree < 100])
net
```

```{r}
# ggnetwork sets default nodes randomly; set seed for reproducibility
set.seed(123)

df_net <- ggnetwork(net, layout = "fruchtermanreingold", weights="Weight", niter=50000, arrow.gap=0)
df_net %>% head()
```

```{r}
plot <- ggplot(df_net, aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(alpha = Weight), size=0.4) +
    geom_nodes(aes(color = as.factor(group), size = centrality)) +
    ggtitle("Network Graph of Breaches from HaveIBeenPwned Database (by @minimaxir)") +
    geom_nodelabel_repel(aes(color = as.factor(group), label = vertex.names),
                          fontface = "bold", size=1.5, box.padding = unit(0.05, "lines"),
                          label.padding= unit(0.1, "lines"), segment.size=0.1, label.size=0.2) +
    scale_alpha_continuous(range=c(0.01,1)) +
    theme_blank() +
    guides(size=FALSE, color=FALSE, alpha=FALSE) +
    theme(plot.title = element_text(family="Source Sans Pro"),
            legend.title = element_text(family="Source Sans Pro"),
            legend.text = element_text(family="Source Sans Pro"))

plot
```

```{r}
ggsave("hibp.png", plot, "png", width=8, height=6, dpi=300)
```

